### 2019-08-07 draft.Rmd
strucutred abstract and language changes

---
title: "A formal approach to child survival from the perspective of mothers"
author: "DA-IW"
date: "August 2019"
output: pdf_document
bibliography: bib.bibtex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Abstract

### Background

An important relation in kinship theory is the expected descendants for mothers at certain age. Child survival and child death affects important dimensions of a mothers' life, such as their reproduction and quality of life.

### Objective

Existing formal demographic tools are used within a stable population framework to characterize child mortality from the perspective of mothers considering both an absolute and a relative change in the force of mortality.

### Results

The expected change in child survival given an absolute change in mortality depends on the difference between maternal age and the age distribution o her surviving children. For absolute changes in mortality, it is necessary to consider a measure of entropy weighted by fertility.

### Contribution

The paper proposes related expressions to describe how mortality changes affects child survival for mothers across demographic regimes. The paper also proposes alternative ways of conceptualizing child survival.

## Relationship

$CS_a$ is the expected number of surviving children to a woman aged $a$ in a stable female population with fertility rates $m_{x}$ and force of mortality $\mu_{x}$:

\begin{equation}
\label{eq:CS}
CS_a = \int_{\alpha}^a{m_{x}l_{a-x}dx}
\end{equation}

We now consider the consequences of a change $\delta$ in mortality in the range $[0,a-\alpha]$, where $\alpha$ is the start age of fertility risk (@goodman1974) and $\delta$ can be absolute or relative (@wrycza2012).

A discrete approximate of the effect on $CS_a$ of an absolute change in mortality $m_{x,\delta}=m_{x}+\delta$ is:

\begin{equation}
\label{eq:abs_change}
\frac{\Delta CS_{a}}{CS_{a}} \approx -(a - \overline{x}_{(CS)}) \Delta\delta.
\end{equation}

In equation \ref{eq:abs_change}, the expected change in $CS_a$ is inversely proportional to the difference between maternal age \textit{a} and the mean age of the mother at the birth of her surviving daughters $\overline{x}_{(CS)}$. In other words, the magnitude of the change depends negatively on the age distribution of the surviving offspring (younger children experience longer periods of exposure to risk).

<!-- No entiendo la oracion de abajo. Le reescribi en el parrafo de arriba, pero no se si esto es lo que querias decir? -->

<!-- In other words, the magnitude of the change depends negatively on how young is the successful reproductive pattern by age, as newborns will have larger risk periods. -->

The relationship is less clear for proportional changes in mortality $m_{x,\delta}=m_{x}(1+\delta)$ given the interaction of birth and mortality rates (@KeyfitzMath). Neverthless we can say that, in discrete terms:

\begin{equation}
\label{eq:relat_change}
\frac{\Delta CS_{a}}{CS_{a}} \approx -H_{CS_a} \Delta\delta.
\end{equation}

Where $H_{CS_a}$ can be interpreted as a temporary entropy constant until age $a-\alpha$ but weighted by the fertility pattern between. In this sense, depending the way in that the survival childs are accumulated by age $C_{CS_a}(x)=\frac{m_{x}l_{a-x}}{CS_{a}}$ is the way that is taken each age contribution to the disparity measure $H_{CS_a}=\int_\alpha^a{C_{CS_a}(x)H(x)}$.
The effect in the child survival experienced by the mother will be bigger when the mortality is the same in all ages, no matter where are concentrated the newborns.

```{r, echo=F}
"graph"
```

Some additional measures are proposed to measure the survival experience by a mother aged $a$. One is an absolute measure of the *Mean Time Spent with Lost*, which can be expressed in terms of a temporary expected lost years measure $e^\dagger$:

\begin{equation}
\label{eq:MTSL}
MTSL_a = \int_\alpha^a{m_x\int_0^{a-x}{d_t e_{t,a-x}dt}dx} = \int_\alpha^a{m_x e_{o,a-x}^\dagger dx}.
\end{equation}

But maybe a mother would relate this lost years with the expected ones, feeling the *intensity* of her losts (*Intensity Time Lost*), and make possible the comparison between fertility regimes:

\begin{equation}
\label{eq:ITL}
  ITL_a = \frac{\int_\alpha^a{m_x e_{o,a-x}^\dagger dx}}{\int_\alpha^a{m_x e_{o,a-x} dx}}
\end{equation}

Finally, different populations could have different mean age at lost of mothers with death childs:

\begin{equation}
\label{eq:MAL}
  MAL_a = K_a + \int_\alpha^a{f_{x,a} MAD_{a-x}}
\end{equation}

$K_a$ is the mean age at birth for womens aged $a$, $MAD_{a-x}$ refers to the mean age at death for those newborns who dies before $a-x$, $TFR_a$ is the cumulated fertility for a women aged a, and $f_{x,a}$ is the cumulative distribution of fertility until age $a$.

## Proofs

### Absolute change

Considering that $m_{x,\delta}=m_{x}+\delta$, $l_{a-x} = e^{-\int_{0}^{a-x}{(\mu_t+\delta})}$, and:

\begin{equation}
CS_{(a)}^\delta = \int_{\alpha}^{a}m_{x} l_{a-x} e^{-\delta (a-x)} dx.  
\end{equation}

To find the effects on daughter survival of adding $\delta$ to the age-specific death rates of daughters, we get the derivative of $dCS_{(a)}^\delta / d\delta$, evaluated near zero (Keyfitz \& Caswell, 2005, section 4.3):

\begin{equation}
\label{eq2}
= -a\int_{\alpha}^{a}m_{x} l_{a-x} dx + \int_{\alpha}^{a}m_{x} l_{a-x}x dx.
\end{equation}

Since $CS_{a} = \int_{\alpha}^{a}m_{x} l_{a-x} dx$, we can rewrite equation (\ref{eq2}) as

\begin{equation}
\frac{dCS^\delta}{d\delta} = -a CS_{a}  + \int_{\alpha}^{a} x m_{x} l_{a-x} dx.
\end{equation}

Dividing both sides by $CS_{a}$ and multiplying by $d\delta$, we get

\begin{equation}
\label{mort_cont}
\frac{d CS_{a}}{CS_{a}} = -(a - \overline{x}_{(CS)}) d\delta
\end{equation}

where $\overline{x}_{(CS)}$ is the mean age of women at the birth of their surviving daughters in a stationary population.

### Relative change

Considering that $m_{x,\delta}=m_{x}(1+\delta)$, $l_{a-x,\delta} = e^{-\int_{0}^{a-x}{\mu_t(1+\delta})dt}=l_{a-x}^{(1+\delta)}$, and:

\begin{equation}
CS_{a}^\delta = \int_{\alpha}^{a} {m_{x} l_{a-x}^{(1+\delta)}} dx.  
\end{equation}

Using that the derivative of $dCS_{(a)}^\delta / d\delta = log(l_{a-x}) l_{a-x}^{(1+\delta)}$, evaluating $\delta$ near zero we get the final expression:

\begin{equation}
\label{mort_change_relative}
\frac{d CS_{a}}{CS_{a}} = \frac{ \int_{\alpha}^{a}m_{x} l_{a-x} \log(l_{a-x}) dx}{\int_{\alpha}^{a}m_{x} l_{a-x}} d\delta
\end{equation}

Using that $C_{CS_a}(x)=\frac{m_{x}l_{a-x}}{CS_{a}}$ is a weighting function and making an analogy with the classical entropy fucntion $H$ where the weights are just the survival function, we can define this inforamtion function $H_{CS_a}=\int_\alpha^a{C_{CS_a}H(x)}$, and get:

\begin{equation}
\label{eq:relat_change2}
\frac{dCS_{a}}{CS_{a}} = -H_{CS_a} d\delta.
\end{equation}

### Mean Age at Lost

Start with the integral of the age $x+t$ at each death child weighted by fertility and survival function

\begin{equation}
\label{eq:MALproof}
\begin{split}  
  MAL_a = \frac{\int_\alpha^a{m_x {\int_0^{a-x}\frac{d_t (x+t) dt}{\int_0^{a-x}{d_tdt}}}}}{\int_\alpha^a{m_x}} \\
  MAL_a = \frac{\int_\alpha^a{m_x x}}{TFR_a} + \frac{\int_\alpha^a{m_x MAD_{a-x}}}{TFR_a} \\
  MAL_a = K_a + \frac{\int_\alpha^a{m_x MAD_{a-x}}}{TFR_a} \\
  MAL_a = K_a + \int_\alpha^a{f_{x,a} MAD_{a-x}}
\end{split}
\end{equation}

## History

Equation  \ref{eq:CS} was developed in @goodman1974, in the "counting method" approach, being the net reproduction rate $R_0$ limited to $a$, or the (not complete) ratio between generations (multiplying and dividing by the mother generation $B$).

@lotka1931orphanhood started to model the orphanhood in a theorical population, comparing the effect in different regimes.  The expected number of living daughters as a stable population concept was mentioned in @goodman1974, extending this idea to the rest of the kinship, in various genealogical directions. The formal effect of different kind of changes in mortality by age was well described in @wrycza2012.

## Applications

The proposed indexes will be applied to two very different regimes.
The changes of mortality will be done computaionally to evaluate which component is more important in the raltive case

## Bibliography