### 2020-04-23 ALAP2020/ALAP.Rmd
* Add visual cheks to approximations and abs change * Add corollary for infant mort abs change * first mature ideas in heterogeneity * graphs must be improved

---
title: 'Child survival for mothers: mortality change and related measures'
author: "Ivan Williams and Diego Alburez-Gutierrez"
date: "07 August 2019"
output:
  html_document:
    fig_caption: yes
bibliography: bib.bibtex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      message=F, warning=F, echo = F)
library(tidyverse)
library(gridExtra)
library(quadprog)
library(wpp2019)
options(digits=2)
source("funs.R")
```

## Introduction

Kin count estimation has a long pedigree in mathematical demography. @lotka1931orphanhood proposed to model the orphanhood in a theorical population, comparing the effect across demographic regimes. Equation  \ref{eq:CS} was originally proposed by @brass_derivation_1953 to estimate child mortality. The notion of estimating the expected number of living daughters in a stable population was generalized by @goodman1974 for other kin relations (granddaughters, cousins, etc.). The "counting method" approach, being the net reproduction rate $R_0$ limited to $a$, or the (not complete) ratio between generations (multiplying and dividing by the mother generation $B$), was further popularized by @KeyfitzMath. @bongaarts_projection_1987 used a similar approach to estimate descendants in his 'Family Status Model'. More recently, @wrycza2012 has looked at the formal effect of different kind of changes in mortality by age.

<!---
Improve Intro: ¿Why is important study survival descendants from mother perspective?¿How impacts mothers health/welfare/experience?
-->

The objective of this work is to give a formal approach to deomgraphy of surviving childs from mother perspective, looking for relations to characterize this situation: the timing of descedent mortality (mean age at the event and time spent in that state) and the effect of mortality change in the final descendants quantity. Numerical applications of the expressions will be done for Latinoamerican countries during the second half of XX Century. Details on data handling and smooth procedures are exposed in [data section](#Data_appendix) in the Appendix.

## Formulas

$CS_a$ is the expected number of surviving children to a mother^[We´ll refer to mothers, assuming that fertility rates applies to all woman, meaning that all of them are exposed.] alive at aged $a$ in a female stable population with fertility rates $m_{x}$, mortality hazard $\mu_{x}$ and survival function $l_x=e^{-\int_{0}^{x}{\mu_t}\,dt}$ ($l_0=1$ during this work), as proposed by @goodman1974:

$$CS_a = \int_{0}^a{m_{x}l_{a-x}dx}$$

Between 1950-1955 and 2010-2015 the portion of surviving daugthers  $\frac{CS_a}{F_a}$ at age 50 passed from a simple mean of `r mean(CS_outs$CS_prob[CS_outs$Period == "1950-1955" & CS_outs$Age==50])` to `r mean(CS_outs$CS_prob[CS_outs$Period == "2010-2015" & CS_outs$Age==50])`, also with a decrease of `r mean(CS_outs$CS[CS_outs$Period == "1950-1955" & CS_outs$Age==50])-mean(CS_outs$CS[CS_outs$Period == "2010-2015" & CS_outs$Age==50])` in the expected descendants at same age. Althought this is not a  real measure but a theorical one if the rates would have been kept constant during enough time.

<!---
Interpretation
-->

```{r plot_CS, fig.cap="Child Survival and Child survival related to cumulated fertility, for women aged 30, 40 and 50. Latinamerican countries in period 1950-2015"}
a <- ggplot(CS_outs, aes(x=Year, y=CS, color=factor(Age)))+
              geom_point(alpha = .3) + 
              geom_smooth(se = F) +
              theme_classic() +
              labs(y = "CS") + theme(legend.position = "none")
b <- ggplot(CS_outs, aes(x=Year, y=CS_prob, color=factor(Age)))+
              geom_point(alpha = .3) + 
              geom_smooth(se = F) +
              theme_classic() +
              labs(y = "CS/F") + 
              scale_colour_discrete(name="Age")
grid.arrange(a,b,ncol=2)
```

### An approximation

Following @KeyfitzMath work about living mothers count, we can consider $a>=\beta$ and assume that fertility pattern by age is concentrated around $\kappa$ (mean age at childbearing), and cumulated fertility (or gross reproduction rate because of female model) is $F_a =  \int_{\alpha}^a{m_{x}dx}$. An approximation of $l_x$ using Taylor´s theorem until second order around $\kappa$ would be:

$$CS_a \approx l_{a-\kappa} \int_{\alpha}^a{m_{x}dx}+(l_{a-\kappa})^{'}\int_{0}^a{(x-\kappa)m_{x}dx}+
(l_{a-\kappa})^{''}\int_{0}^a{\frac{(x-\kappa)^2}{2} m_{x}dx}\\
  \approx F_a l_{a-\kappa} + \frac{\sigma^2}{2}F_a (l_{a-\kappa})^{''}\\
  \approx F_a \, l_{a-\kappa}[1+\frac{\sigma^2}{2}\frac{(l_{a-\kappa})^{''}}{l_{a-\kappa}}]$$

The second Taylor´s term is null because $\int_{\alpha}^a{x \, m_{x}dx} = \kappa \, F_a$. The formula tells us that the survival experience for a woman aged *a* depends mainly of cumulated fertility (possible cases, different from the ancestor count case where this doesn´t affect) and child survival probability (success cases), but also negative proportional affected by how disperse is the fertility by age (variance \sigma^2) and the suvival curvature around (negative sign), in a $l_x$ range where is tipycally very flat (range 20-40 years old) in transitioned populations. Guatemala improved the approximation with years due to rectangularization process in $l_x$ (figure \@ref(fig:plot_CS_aprox)).

```{r plot_CS_aprox, fig.cap = "Error in approximation for a=40, from years 1950 (darker) to 2015 (lighter). Guatemala in red (left). Guatemala survival and fertility by age (right)"}
a <- ggplot(CS_outs %>% filter(Age == 40), 
                    aes(x=CS, y=error_aprox, color=Year))+
                    geom_point(alpha = .6) + 
                    geom_point(data = CS_outs %>% filter(name=="Guatemala",Age == 50),
                               aes(x=CS, y=error_aprox), color="red") +
                    theme_classic() +
                    theme(legend.position = "none") +
                    labs(y = "%")
b <- LA %>% filter(name=="Guatemala") %>% ggplot() +
            geom_line(aes(x=x, y=lx/100000, color=Period)) +
                    theme_classic() +
                    scale_color_grey()+
                    theme(legend.position = "none") +
                    labs(y = "")
c <- LA %>% filter(name=="Guatemala", between(x, 15,50)) %>% ggplot() +
            geom_line(aes(x=x, y=asfr/tfr*100, color=Period)) +
                    theme_classic() +
                    scale_color_grey()+
                    theme(legend.position = "none") +
                    labs(y = "%")
grid.arrange(a, grid.arrange(b,c,ncol=1), ncol=2)
```

### Burden and timing

We call *Mean Time Spent in Bereavement* (*MTSB*) the absolute measure of the expected total lifetime with a death daughter for a mother aged *a*, which can be expressed in terms of a temporary expected lost years index in line with $e^\dagger$ (@Vaupel1986):

$$MTSB_a = \int_0^a{m_x\int_0^{a-x}{d_t e_{0|a-x-t}\, dt}\,dx} = \int_\alpha^a{m_x e_{0|a-x}^\dagger \,dx}$$.

Where $d_t$ is the death distribution from birth, $e_{0|a-x-t}$ is the life expectancy at birth until age $a-x-t$ and $e_{0|a-x}^\dagger$ the temporary dispertion measure. But most interesting would be to compare these years with the time that mothers would expect to live with their daughters. We call this the *Intensity Time in Bereavement* ($ITB$)): a ratio between expected time with a "lost" and expected time with a "life", that allows to make comparisons between population regimes.

$$ITL_a = \frac{\int_0^a{m_x e_{0|{a-x}}^\dagger dx}}{\int_0^a{m_x e_{0|a-x} dx}}$$

This is a ratio between child-years in two radically different states. Looks similar to the trascendental entropy measure $H$ (@KeyfitzMath) but considering all the cohorts born during the the mother´s life, weighted by their relative size $m_x$. In the figure \@ref(fig:plot_ITL_MAL) is shown that *ITB* is bigger for youg womans because of the weight of infant mortality in their mother experience, also with more dispersion for same levels of parity. Assuming constant period rates the expected time in years that a mother aged 30 passed with a death son was around 4% in some countries at middle XX Century. When increasing age, the survival experience depends less on infant mortality, and the distribution is around less than 2% for womans aged 50, converging to 0 on time. As extreme cases, in 1950-1955 Haiti women aged 30 would have experienced an intensity of `r round(CS_outs$ITL[CS_outs$name=="Haiti" & CS_outs$Age==30 & CS_outs$Period=="1950-1955"],1)`%, and in 2010-2015 `r round(CS_outs$ITL[CS_outs$name=="Haiti" &CS_outs$Age==30 & CS_outs$Period=="2010-2015"],1)`%, while the women of Costa Rica `r round(CS_outs$ITL[CS_outs$name=="Costa Rica" & CS_outs$Age==30 & CS_outs$Period=="1950-1955"],1)`%, and in 2010-2015 `r round(CS_outs$ITL[CS_outs$name=="Costa Rica" &CS_outs$Age==30 & CS_outs$Period=="2010-2015"],1)`%.

Another important factor for the child survival experience of mothers, is the mean age at child loss, called here $MAL$. This relation can be derived by starting with the mother age $x+t$ at each death child age $t$ at death, weighted by the fertility and survival function. In it, $\kappa$ is the mean age at childbirth for womens aged $a$, $MAD_{a-x}$ refers to the mean age at death for newborns that die before $a-x$, $F_a$ is the cumulated fertility for a women aged $a$:

$$MAL_a = \frac{\int_0^a{m_x {\frac{\int_0^{a-x} l_t \, \mu_t (x+t) dt}{\int_0^{a-x} l_t \, \mu_t dt}}dx}}{\int_0^a{m_x}dx} \\
  MAL_a = \frac{\int_0^a{m_x \left[{x+\frac{\int_0^{a-x} l_t \, \mu_t \, t \, dt}{\int_0^{a-x} l_t \, \mu_t dt}}\right]dx}}{\int_0^a{m_x}} \\
  MAL_a = \frac{\int_0^a{m_x \,x \,dx}}{F_a} + \frac{\int_0^a{m_x MAD_{a-x}dx}}{F_a} \\
  MAL_a = \kappa + \frac{\int_0^a{m_x MAD_{a-x}dx}}{F_a}$$

For those latinamerican mothers aged 50 with 3 daughters born who suffered a lost, theyy experienced that at age 30 in average (figure \@ref(fig:plot_ITL_MAL)) (More can be analyzed here).

```{r plot_ITL_MAL, fig.cap = "Intensity Tome Lost of  womans aged 30, 40 and 50. Years 1950 (darker) to 2015 (lighter). Latinamerican countries in period 1950-2015"}
ITL <- ggplot(CS_outs)+  
                    geom_smooth(method="lm", aes(x=Fa, y=ITL, color=factor(Age)), 
                                se = F, formula= (y ~ exp(x)))+
                    geom_point(aes(x=Fa, y=ITL, color=factor(Age)), alpha = .3) +
                    theme_classic()+        
                    theme(legend.position = "right") +
                    labs(y = "%")

MAL <- ggplot(CS_outs %>% filter(Age==50))+
                    geom_smooth(aes(x=Fa, y=MAL), method = "lm", col="grey", se = F)+
                    geom_point(aes(x=Fa, y=MAL, color=Year), alpha = .3) +
                    theme_classic()+
                    theme(legend.position = "none") +
                    labs(y = "Mean Age at Lost")
grid.arrange(ITL, MAL, ncol=2)
```

### Mortality changes

We now consider the consequences of an absolute change $\delta$ in mortality in the range $[0,a-\alpha]$, where $\alpha$ is the start age of fertility risk. Considering that $m_{x,\delta}=m_{x}+\delta$ (@wrycza2012), and $l_{a-x}^\delta = e^{-\int_{0}^{a-x}{(\mu_t+\delta})dt}$:

$$CS_{a}^\delta = \int_{\alpha}^{a}m_{x} l_{a-x} e^{-\delta (a-x)} dx$$

We get the derivative of $dCS_{(a)}^\delta / d\delta$ evaluated near zero (@KeyfitzMath) to find the effects of adding $\delta$ to hazard rates:

$$\frac{dCS^{\delta}}{d\delta} = -a\int_{\alpha}^{a}m_{x}{l_{a-x} dx} + \int_{\alpha}^{a} m_{x} {l_{a-x}x dx}\\
= -a \, CS_{a}  + \int_{\alpha}^{a} {x  \, m_{x} l_{a-x} dx}$$

Dividing both sides by $CS_{a}$ in an discrete approzimation we get

$$\frac{\Delta CS_{a}}{CS_{a}} \approx -(a - k_a) \Delta\delta$$

The expected change in descendants survival is inversely proportional to the difference between maternal age $a$ and the mean age of the mother at the birth of her surviving daughters $k_a$ (always smaller than $\kappa_a$). The magnitude of the change depends negatively on the age distribution of the surviving offspring. This is intuitive considering that younger descendants experiences longer periods of exposure to risk. The figure \@ref(fig:CS_abs_app) shows that goodness of fit is descreasing with the change size, given that $\delta$ is assumed near zero (as an example considering $\delta=0.01$ means at age 50 that $\frac{l_50}{e^{\delta\,50}} = \frac{l_50}{1.6}$, a big change).

```{r CS_abs_app, fig.cap= "Absolute change and its approximation for age CS in women aged 50. Latinamerican countries in period 1950-2015"}
# application of absolute changes
age = 50
CS_abs_change_app = data.frame(delta = NA,Country = NA, Period = NA, change = 0, change_app = 0)
for(delta in seq(0,.01,.001)){
  for (country in unique(LA$name)){
    for (period in unique(LA$Period)){
      # country = "Argentina"; period = "2010-2015"; delta = .01
      change_app = -(age - CS_outs %>% filter(name == country, 
                                              Period == period, 
                                              Age == age) %>% select(k)) * delta
      country_i = LA %>% filter(name == country, Period == period) 
      CS = CS_outs %>% filter(name == country, 
                              Period == period, 
                              Age == age) %>% select(CS)
      CS_changed = sum(unlist(sapply(1:age, function(i) 
        country_i$asfr[i]*country_i$lx[age-(i-1)]/100000*exp(-delta*(age-(i-1))))), 
        na.rm = T)
      change = (CS_changed-CS)/CS
      CS_abs_change_app = rbind(CS_abs_change_app, 
                                data.frame(delta = delta,
                                           Country = country, Period = period, 
                                           change = as.numeric(change), 
                                           change_app = as.numeric(change_app)))
    }
  } 
}
CS_abs_change_app %>% slice(-1) %>% 
  ggplot() + 
  geom_point(aes(x=change, y=change_app, color=delta)) +
  geom_abline(slope=1, intercept = 0) +
  theme_classic() +
  # theme(legend.position = "none")+
  xlim(-.3, 0) + ylim(-.3, 0)
```

That an equal change in mortality could happens at all ages is unlikelihood. As a [corollary](#Corollary) we inspect the case thtat only infant mortality is affected.

The relationship is less clear for a proportional change given the interaction of birth and mortality rates (@KeyfitzMath). Considering that $\mu_{x,\delta}=\mu_{x}(1+\delta)$ and $l_{a-x}^\delta = e^{-\int_{0}^{a-x}{\mu_t(1+\delta})dt}=(l_{a-x})^{(1+\delta)}$:

$$CS_{a}^\delta = \int_{\alpha}^{a} {m_{x} l_{a-x}^{(1+\delta)}} dx$$

Using the derivative $\frac{dl_{a-x}^{(1+\delta)}}{d\delta} = log(l_{a-x}) l_{a-x}^{(1+\delta)}$, and in the third row reversing integrals between $t$ and $x$:

$$\frac{d CS_{a}^\delta}{d \delta} = \int_{\alpha}^{a} {m_{x} l_{a-x} \log(l_{a-x}) dx}\\
= - \int_{0}^{a} {m_{x} l_{a-x} \int_{0}^{a-x}{\mu_x \, dt}\, dx}\\
= - \int_{0}^{a}{\mu_t} \int_{0}^{a-t} {m_{x} l_{a-t-x} \, dx}\, dt$$

Dividing by $CS_a$ and multiplying by $\delta$ get one useful expression. This is the negative of cumulative hazard $H_a$ but considering a factor $0\leq \frac{CS_{a-t}}{CS_a}\leq1$ that takes in acccount the amount of survinvig descendatns that would be lost for a mother aged *a* in each age x of risk.

$$\frac{\Delta CS_{a}}{CS_a} = - \left[\int_{0}^{a}{\mu_t  \, \frac{CS_{a-t}}{CS_a} \,dt \,}\right] \Delta \delta$$

This factor has a S-shape because $l_x$ curvature fertility acumulation, that gives more weight to first ages, and which is grater than 1 only for ages *a* with complete reproductive life and high infant mortality (with no aditional childs and higher risk for one age additional of survive).

<!---
An Application here
-->

```{r chek_CS_ratio, include = F}
# test factor
countries = c("Guatemala", "Argentina", "Haiti")
period = "1950-1955"
Age = 40
plot(NULL, xlim=c(15,Age), ylim=c(0,1.1),xlab="Age", ylab="Ratio")
abline(h=1, lty=2, col="grey")
for (country in countries){
  points(15:Age,
         sapply(15:Age, function(age)
         CS_results(country = country, 
             period = period, 
             a = age)$CS)/
          CS_results(country = country, 
             period = period, 
             a = Age)$CS,
          col = which(country==countries), t="l")
}
```

### Heterogeneity

In a model with heterogeneity, determined at born with a multiplicative effect, $CS_a$ could be interpreted as a conditional expectation with random variables *K* for fertility heterogeneity and *Z* for mortality frailty.

$$CS_a(k,z) = \int_{0}^{\infty}\,{m(k)}_x\,{l(z)}_{a-x}\,dx$$

Following @Coresh1988 we can express the fertility part in a multiplicative way with variability in the level but not in the shape, as ${m(k)}_x = F_\beta\,k\,r_x$, where $F_\beta$ is the baseline cumulated fertility, $K$ is a random variable with mean 1 that allows variability between groups and $r_x$ is the fertility structure by age ($\int_{0}^{\beta}{r_x}=1$). Frailty part can be thought in a cohort effect way, also in the multiplicative assumption as @Vaupel2014, with $l(z)_x=e^{-H_{x}Z}=l_x^Z$, with baseline hazard $\mu$. Replacing conveniently the last two expressions and considering the joint distribution $f_{kz}=f_{z/k} \,f_k$, the mean $\overline{CS}_a$ would be:

$$\overline{CS}_a = \int_{0}^{\infty}\int_{0}^{\infty}\left[\int_{0}^{a}{ {m(k)}_x l(z)_{a-x}\,dx}\right]f_{kz}\,dz \,dk\\
= F_\beta \int_{0}^{a-x} r_x \int_{0}^{\infty} \int_{0}^{\infty} {l_{a-x}}^z \, k\, f_{kz} \,dk\,dz\,dx$$

The way to chain dependence is convenient to the expected herarchy $f_{zk}=f_{k|z}f_z$: groups with higher descendant´s mortality would adjust their fertility level, with positive correlation. Isolating one age and defining $f_{kz}$ we can explore effects in $CS_a$.

<!---
More development here
-->

```{r, include=F}
# frailty distributions
db = LA %>% filter(x==10) %>% select(q,tfr)
cor.test(LA$q[LA$x==10],LA$tfr[LA$x==10])
plot(db)
model = lm(tfr~log(q), db)
lines(seq(0,0.4,0.01), predict(model,newdata = data.frame(q=seq(0,0.4,0.01))))
```

## Discussion

## Appendix

### Data {#Data_appendix}

Mortality and fertility data was obtained from the last revision of population prospects by UN (@WPP19). We smoothed female $l_x$ using cubic-splines constrained to monothonic descense, taking later $L_0$ and $T_{100}$ as inputs for year-person calculations. For splitting fertility five groups was used quadratic optimization approach by @Michalski2018, with an desirable property for our purpose which is a good fitting in parity. Also was assumed an unique female percentage of newborns of `r round(im, 4)` for all period-country cases.

An negative relation between mortality until 10 and TFR is shown in \@ref(fig:plot_tfr_q010). Specific paths like Argentina and Guatemala shows very different demographic profiles during second part of XX century.

```{r plot_tfr_q010, fig.cap="Female probability of death in first 10 years old by gross reproduction rate. Latin America countries in period 1950-2015"}
ggplot(LA %>% filter(x==10), aes(x=tfr * 2.04, y=q))+
                  geom_point(aes(color=Period)) + 
                  geom_smooth(se = F, color = "grey", method = "lm", 
                              formula = y ~ exp(x)) +
                  geom_line(data = LA %>% filter(x==10, name %in% c("Guatemala")),
                            aes(group=name), colour=2, size = 1) +
                  geom_line(data = LA %>% filter(x==10, name %in% c("Argentina")),
                            aes(group=name), colour=4, size = 1) +
                  theme_classic() + labs(x="GRR", y="q(0,10)")
```

### Corollary {#Corollary}

That an equal change in mortality could happens at all ages is unlikelihood. This measure is much more related to first ages (child and youth) than adult or ones. As a corollary, if the change only happens in the age range $\left[0;1\right)$, infant mortality, we can inspect the effect starting with splitting the integral (of course, for $a>1$)]:

$$CS_{a}^{\delta_0} = \int_{0}^{a-1}
m_{x} {e{^{-\int_{1}^{a-x}\mu_t \,dt-\int_{0}^{1}(\mu_t + \delta)\,dt}}} dx + 
\int_{a-1}^{a} m_{x} e^{-\int_{0}^{a-x}(\mu_t+ \delta)\,dt}\\
=\int_{0}^{a-1} m_{x} l_{a-x}e^{-\delta} dx + \int_{a-1}^{a} m_{x} l_{a-x} e^{-\delta(a-x)}\,dx$$

Deriving by $\delta$ and valuating near 0, we get:

$$CS_{a}^{\delta_0} =-\int_{0}^{a-1} m_{x} l_{a-x} dx - a \int_{a-1}^{a} {m_{x} l_{a-x} dx} + \int_{a-1}^{a} {x \, m_{x} l_{a-x} dx}$$
If we name $CS_{a-1,a}=\int_{a-1}^{a} {m_{x} l_{a-x} dx}$, we can express the first term in the right as $CS_a - CS_{a-1,a}$, and finally divide by $CS_a$: 
$$ = - CS_a + CS_{a-1,a} (1-a+\kappa_{a-1,a})\\
\frac{\Delta CS_{a}}{CS_{a}} \approx - \left[1-\frac{CS_{a-1,a}}{CS_a} (1-(a-\kappa_{a-1,a}))\right] \Delta\delta$$

This means that an absolute change in infant hazard rate affects proportionally for all the age range extension (because of all the cohorts aged more than 1), except for the mean time exposure for those who born and survive in the last year $1-(a-k_{a-1,a})$, times $\frac{CS_{a-1,a}}{CS_a}$ the portion of cumulated alive daughters for the same period, weighting how important was this in total succesful experience (in terms of alive descendants).

## References