### 2020-05-14 covid/covid_proposal.Rmd

---
title: 'Estimating excess bereavement due to Covid-19'
subtitle: 'A research proposal'
author: "Diego Alburez-Gutierrez and Ivan Williams"
date: "15 May 2019"
output:
  pdf_document:
    fig_caption: yes
bibliography: bib.bibtex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      message=F, warning=F, echo = F)
```

## Introduction

Death and bereavement has come to the forefront of public debate as the world grapples with the global pandemic of the Covid-19 disease, which has already brought about the death of thousands of mainly elderly people. 
Each death represents the loss of somebody's parent, grandparent, great-grandparent, aunt, uncle, cousin, or child.
Studies in sociology and public health have consistently shown the negative and long-term consequences of bereavement on mental, physical, and emotional health, especially for women [@umberson_death_2017].
Bereavement also matters because relatives are crucial providers of social and financial support and their absence will be direly felt by bereaved relatives [@Hendrickson2009].
For many people living in the Global North this will be their first close encouter with death, as historically low mortality rates have implied a reduction in the exposure to mortality at all ages in high-income countries [@murphy_long-term_2011; @alburez_gutierrez_womens_2019]. 
For people in the Global South, it will add to an already high burden of bereavement [@smith-greenaway_maternal_2020; @roberton_early_2020].

Much attention has been given to Covid-19 excess mortality rates, but little is known about how the pandemic will increase the expossure to death for people around the world.
This project will combine existing and future data on Covid-19 excess mortality with powerful methods from formal demography [@goodman1974] to estimate the burden of Covid-19 bereavement.
The project is not interested in the total number of people who will die from Covid-19, but in the number of people who will be affected by the death of a relative because of the disease.
We are interested in `excess bereavement'.
We ask:

   1. How many people will lose a relative to Covid-19, either in the nuclear or in the extended family? 
   2. What is the age distribution of the bereaved population?
   3. How will the experience of bereavement affect the availability of kin resources for those that survive the disease?

## Project description

### Methodological component

We seek to generalize the Kin-Cohort Method [@alburez_gutierrez_womens_2019], itself an extension of the Goodman, Keyfitz Pullum kinship equations [@goodman1974], to estimate the potential increase in the expossure to the death of relatives. 
A first step will focus on deriving mathematical expressions to model the effect of changes in mortality on the population-level prevalence of bereavement.

We exemplify this for the case of parental bereavement, which we understand as the probability of losing a child. 
Let $CS_a$ be the expected number of surviving children to a mother alive at age $a$ in a female stable population with fertility rates $m_{x}$, mortality hazard $\mu_{x}$ and survival function $l_x=e^{-\int_{0}^{x}{\mu_t}\,dt}$ (with unit radix $l_0=1$), as proposed by @goodman1974:

$$CS_a = \int_{0}^a{m_{x}l_{a-x}dx}.$$

Now consider an absolute increase in mortality $\delta$ in mortality in the range $[0,a-\alpha]$, where $\alpha$ is the start age of fertility risk.
We can think of this hypothetical increase in mortality as the excess mortality derived from the Covid-19 disease so that $m_{x,\delta}=m_{x}+\delta$ (@wrycza2012) and $l_{a-x}^\delta = e^{-\int_{0}^{a-x}{(\mu_t+\delta})dt}$:

$$CS_{a}^\delta = \int_{\alpha}^{a}m_{x} l_{a-x} e^{-\delta (a-x)} dx.$$

As we have shown elsewhere [ref], we can conceptualize the effects of this absolute change $\delta$ in mortality as the difference between maternal age $a$ and the mean age of the mother at the birth of her surviving daughters $k_a$:

<!-- Si sacamos nuestro paper como pre-print, lo podemos citar aqui!! -->

$$\frac{\Delta CS_{a}}{CS_{a}} \approx -(a - k_a) \Delta\delta.$$

Deriving these formal relationships for other types of relative (mothers, grandmothers, siblings, cousins, and aunts) will provide key insight for the empirical analysis of excess bereavement. 
Further development is needed to consider age-specific changes in mortality, as the Covid-19 disease affects mainly people in older ages.
Finally, the existing models ignore heterogeneity in the experience of kin death and within-group correlations.
This is a limitation considering that deaths from infectious disease are likely to be clustered in the population.
The final strand of the methodological component of the project will adress these issues.

### Empirical component

We will estimate the excess burden of kin death from the Covid-19 disease using our novel mathematical models with mortality and fertility rates as inputs.
Combining this with data on the age and sex structure of the population will allow us to estimate the magnitude of the excess bereavement and the age distribution of the bereaved population\footnote{`Excess' bereavement can be compared to a baseline counter-factual scenario using rates that exclude the excess mortality from Covid-19.}.

As an initial step, the analysis will focus on Germany and Italy, for which quality data on Covid-19 death rates is already available.
Further development may apply our innovative methodology to other setting once data on Covid-19 excess mortality becomes available. 
The project will produce a range of estimates, given the uncertainty inherent to the Covid-19 mortality statistics that are required as input. 
However, we expect the accuracy of our models to improve as more data quality becomes available.

## Expected main results

- A flexible and scalable methodology for estimating the excess burden of covid-19 bereavement. The open-source code to produce the estimates will be made freely available on the code-sharing platform [Github](www.github.com).
- The first set of estimates of how the Covid-19 disease will affect the expossure to mortality at a population-level
- A set of academic publications describing the methodology and the main results of the project
- A online platform for scientists and policy makers to access and download the data easily (see [this interactive app](https://research-app.shinyapps.io/child_death_paa/))

<!-- DIAGRAMA DE PARENTEZCO, DONDE EL EFECTO DE MORTALITYDAD ES LA FLECHA Y EL NODO EL NUMERO ESPERADO DE FAMILIARES -->

## References