### 2020-06-02 covid/latest/attachments/covid-19 VW grant/CV_Diego_Alburez_VW.tex
added cvs

% LaTeX file for resume 
% This file uses the resume document class (res.cls)

\documentclass{res} 
%\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
%\usepackage{newcent}   % uses new century schoolbook postscript font 
\newsectionwidth{0pt}  % So the text is not indented under section headings
\usepackage{fancyhdr}  % use this package to get a 2 line header
\usepackage{booktabs}
\usepackage{hyperref}
%\usepackage{tabularx}
\usepackage{lipsum}

\hypersetup{
    colorlinks=true,
    urlcolor=blue,
}

\renewcommand{\headrulewidth}{0pt} % suppress line drawn by default by fancyhdr
\setlength{\headheight}{24pt} % allow room for 2-line header
\setlength{\headsep}{24pt}  % space between header and text
\setlength{\headheight}{24pt} % allow room for 2-line header
\pagestyle{fancy}     % set pagestyle for document
\rhead{ {\it D. Alburez}\\{\it p. \thepage} } % put text in header (right side)
\cfoot{}                                     % the foot is empty
\topmargin=-0.5in % start text higher on the page

\begin{document}
\thispagestyle{empty} % this page has no header

\today

\name{Diego Alburez-Gutierrez\\[12pt]}% the \\[12pt] adds a blank line after name

\address{Konrad-Zuse-Stra{\ss}e 1, 18057 Rostock, Germany \\  \href{mailto:alburezgutierrez@demogr.mpg.de}{alburezgutierrez@demogr.mpg.de}; \href{www.alburez.me}{www.alburez.me} }

% \href{https://twitter.com/d_alburez}{@d\textunderscore alburez}

\begin{resume}

I am a social scientist specialising in inter-generational demography. 
My work has used mathematical modeling, micro-simulation techniques, and empirical analysis to explore changes in the experience of life events such as kin availability, family bereavement, and caring responsibilities across demographic regimes.

%\vspace{0.2in}
\section{\centerline{EDUCATION}}

\textbf{The London School of Economics and Political Science, UK} \\
{\sl PhD}, Demography and Population Studies \hfill 2014 - 2018

{\sl MSc},  Social Research Methods (Distinction) \hfill 
 2013-2014

\textbf{Del Valle University, Guatemala} \\
{\sl BA},  Anthropology (First class) \hfill 
2008-2012

%\vspace{0.2in}

\section{\centerline{AWARDS AND SCHOLARSHIPS}}

   \begin{itemize}
    \item 2019 \href{https://www.mpg.de/prizes/otto-hahn-medal}{Otto Hahn Medal} (2020) awarded by the Max Planck Society
   	\item LSE PhD Academy Prize (2018) awarded by the London School of Economics for research on state-sponsored massacres in Guatemala
 	\item LSE PhD Studentship (2014-2018) awarded by the London School of Economics
%    \item \href{https://www.mpg.de/prizes/otto-hahn-medal}{Otto Hahn Medal} (2019; EUR 7,000 $\approx$ USD 7,500) awarded by the Max Planck Society for `outstanding achievement from the initial creative period of young scientists', in particular `for methodological innovations for studying the long-term consequences of mass killings on local populations' -
%   	\item LSE PhD Academy Prize (2018) for research on state-sponsored massacres in Guatemala
% 	\item LSE PhD Studentship (2014-2018; GBP 72,000 $\approx$ USD 121,000) 
 \end{itemize}

\section{\centerline{RESEARCH EXPERIENCE}}

\textbf{Max Planck Institute for Demographic Research (MPIDR), Germany} \\
{\sl Research Scientist (post-doc) and Research Area Chair (Aging and Generational Processes)}, Lab of Digital and Computational Demography \hfill        Since 2018

\textbf{United Nations High Commissioner for Refugees (UNHCR), Denmark} \\
{\sl Statistics Team Intern}, Field Information and Coordination Support Section (FICSS) \hfill        2018

\textbf{The London School of Economics and Political Science, UK} \\
{\sl Research Assistant}, positions in the Sociology, Social Policy, Methodology, and Law departments  \hfill       2013-2017

\textbf{The Population Council}, Guatemala \\
{\sl Research Sub Coordinator} \hfill    2012

\section{\centerline{PUBLICATIONS IN PEER-REVIEWED JOURNALS}}

\begin{flushleft}
\hangindent=2em
Nepomuceno, M., Acosta, E., \textbf{Alburez-Gutierrez, D.}, Aburto, J.M., Gagnon, A., and Turra, C. (Forthcoming). ``Besides population age-structure, health and other demographic factors can contribute to understanding the COVID-19 burden across the globe." Proceedings of the National Academy of Sciences (Accepted for publication). \\
\textbf{Alburez-Gutierrez, D.}, Chandrasekharan, E., Chunara, R., Gil-Clavel, S., Hannak, A., Interdonato, R., Joseph, K., Kalimeri, K., Malik, M., Mayer, K., Mejova, Y., Paolotti, D., and Zagheni, E. (2019). ``Reports of the Workshops Held at the 2019 International AAAI Conference on Web and Social Media." AI Magazine, 40(4), 78-82. doi:\href{https://doi.org/10.1609/aimag.v40i4.5287}{10.1609/aimag.v40i4.5287}. \\
Leone, T., \textbf{Alburez-Gutierrez, D.}, Gandour, R., Coast, E., and Giacaman, R. (2019). ``Maternal and child access to care and intensity of conflict in the occupied Palestinian territory: a pseudo-longitudinal analysis (2000–2014)." Conflict and Health 13(1). doi:\href{https://doi.org/10.1186/s13031-019-0220-2}{10.1186/s13031-019-0220-2}. \\
\textbf{Alburez-Gutierrez, D}. (2019). Blood is thicker than bloodshed: A genealogical approach to reconstruct populations after armed conflicts. Demographic Research, 40(23): 627-656. doi:\href{https://www.demographic-research.org/volumes/vol40/23/}{10.4054/DemRes.2019.40.23}. \\
McGovern, P., and \textbf{D. Alburez-Gutierrez}. (2017) ``Who takes workplace case-study methods seriously? The influence of gender, academic rank and PhD training." Industrial Relations Journal 48(2):98-114, doi:\href{http://onlinelibrary.wiley.com/doi/10.1111/irj.12171/full}{10.1111/irj.12171}.
\end{flushleft}

\section{\centerline{BOOK CHAPTERS AND TECHNICAL REPORTS}}

\begin{flushleft}

\hangindent=2em
\textbf{Alburez-Gutierrez, D.}, S. Aref, S. Gil-Clavel, A. Grow, D.V. Negraia, E. Zagheni. (2019). ``Demography in the Digital Era: New Data Sources for Population Research.'' Proceedings of the 2019 Conference of the Italian Statistical Society. doi: \href{https://doi.org/10.31235/osf.io/24jp7}{10.31235/osf.io/24jp7}. \\
\hangindent=2em
\textbf{Alburez-Gutierrez, D.} and Segura Garcia, C. (2018). ``The UNHCR Demographic Projection Tool: estimating the future size and composition of forcibly displaced populations." UNHCR Statistics Technical Series, \href{http://www.unhcr.org/en-au/5ae9ee747.pdf}{2018/1}.

\end{flushleft}

\section{\centerline{INVITED PRESENTATIONS (THREE SELECTED)}}

\begin{flushleft}

\hangindent=2em
\textbf{Alburez-Gutierrez, D.} (27 May 2020, postponed due to Covid-19 lockdown) ``Die demografischen Auswirkungen von Massengewalt in Guatemala [The demographic consequences of Genocide in Guatemala]." Berlin, Germany (Ibero-American Institute). \\
\hangindent=2em
\textbf{Alburez-Gutierrez, D.} (27 Nov 2019) ``Women's experience of child death: a global demographic perspective." \href{https://www.iser.essex.ac.uk/events/seminars/external/2019-11-27}{ISER External Seminar}, Colchester, UK (University of Essex). \\
\hangindent=2em
\textbf{Alburez-Gutierrez, D.} (23 May 2019) ``The Demographic Foundations of the Lived Experience of Kin Death: A Genealogical Reconstruction." \href{https://www.suda.su.se/about-us/events/suda-demographic-colloquium-series/suda-demographic-colloquium-diego-alburez-gutierrez-max-planck-institute-for-demographic-research-1.423293}{SUDA Demographic Colloquium Series}, Stockholm University.

\end{flushleft}

\section{\centerline{OUTREACH AND SCIENCE COMMUNICATION}}

\begin{itemize}
  \item Academic video: `How Can We Quantify the Immediate and Lingering Impact of Genocide on Local Populations?' (\href{https://lt.org/publication/how-can-we-quantify-immediate-and-lingering-impact-genocide-local-populations}{Latest Thinking, Apr 2020})
  \item Opinion piece: `Coronavirus: a warning to Latin America and the Caribbean to dramatically increase COVID-19 testing' (\href{https://theconversation.com/coronavirus-a-warning-to-latin-america-and-the-caribbean-to-dramatically-increase-covid-19-testing-135759}{The Conversation, UK, Apr 2020})
  \item Blog post: `Genocide is the tip of the iceberg: Reviewing the Guatemalan case' (\href{https://blogs.lse.ac.uk/humanrights/2016/02/18/genocide-is-the-tip-of-the-iceberg-reviewing-the-guatemalan-case/?fbclid=IwAR3mPVASOpmBfAu19BMXk99_TGgo6-0-axr6ZT0i5msU7atkppm4rgant_Q}{LSE Human Rights Blog, Feb 2016})
\end{itemize}

\section{\centerline{ORGANISER OF INTERNATIONAL ACADEMIC EVENTS}}

\begin{itemize}
    \item \textit{Organiser}: \href{https://www.demogr.mpg.de/en/news_events_6123/news_press_releases_4630/news/open_access_to_demographic_research_6699}{Rostock Open Science Workshop}, Max Planck Institute for Demographic Research (MPIDR), Rostock\hfill 2019
    \item \textit{Organiser}: \href{https://iussp.org/en/workshop-demographic-research-web-and-social-media-data}{Workshop on Demographic Research with Web and Social Media Data}, $13^{th}$ International AAAI Conference on Web and Social Media (ICWSM), Munich \hfill 2019
\end{itemize}

\section{\centerline{SERVICE TO THE DISCIPLINE}}

\begin{itemize}
    \item \textit{Reviewer} for academic journals: Demography; Journal of Marriage and Family; Demographic Research; Revue Quetelet/Quetelet Journal; $11^{th}$ International Conference on Social Informatics (SocInfo `19)
	\item \textit{Organiser} of the `Social Policy Qualitative Research Group', London School of Economics  \hfill 2017-2018 
	\item \textit{President} of the `Methodology Student Union', London School of Economics \hfill 2013-2014
%	\item \textit{Organiser} of the II Colloquium of Anthropology and Sociology, Universidad Del Valle\hfill 2011 
\end{itemize}

\end{resume} 
\end{document}